How many researchers use 'boilerplate' statistical analysis sections?

### 2021-03-12 manuscript/draft_manuscript.Rmd
Update draft_manuscript.Rmd

_PLOS ONE_ is a large open access journal that publishes original research across a wide range of scientific fields. Article submissions are handled by an academic editor who selects peer reviewers based on their self-nominated area(s) of expertise. Submissions do not undergo formal statistical review. Instead, reviewers are required to assess submissions against several publication criteria, including whether: “Experiments, statistics, and other analyses are performed to a high technical standard and are described in sufficient detail”. Authors are encouraged to follow published reporting guidelines to ensure that chosen statistical methods are appropriate for the study design, and adequate details are provided to enable independent replication of results. Reporting guidelines are available from the EQUATOR network which has developed guidelines for study designs commonly used in health research, and these guidelines were developed to tackle poor statistical application and reporting in health and medical journals [@Altman2016].

@@

_Step 1_: Targeted API searches. API searches were completed using the R package ‘rplos’ [@rplos]. Search queries targeted the presence of analysis-related terms anywhere in the article. Search terms combined the words “data” or “statistical” with one of: “analysis”, “analyses”, “method”, “methodology” or “model(l)ing”. Search terms were intended to be broad whilst keeping search results to a manageable number for full-text review (see Step 2). By allowing terms to appear anywhere within the article, we accounted for the possibility of relevant text being placed in different sections, for example, in the Material and Methods section versus Supplementary Information. Search results were indexed by a unique Digital Object Identifier (DOI). Attribute data collected per DOI included journal volume, subject classification(s) and total article views since publication.

@@

The ANZCTR was established in 2005 as part of a coordinated global effort to improve research quality and transparency in clinical trials reporting; observational studies can also be registered. All studies registered on ANZCTR are publicly available and can be searched for via an online portal. 
<!--- Researchers responsible for overseeing a clinical trial are expected to register full details before commencing participant enrolment. --->
Details required for registration follow a standardised template (reference or supp file), which covers participant eligibility, intervention(s) being evaluated, study design and outcomes. Researchers are asked to provide a "brief description" of the sample size calculations, statistical methods and planned analyses, although this section is not compulsory [@ANZCTR]. 
<!---Upon successful registration, the trial is assigned a unique registration number for use in future reporting in scientific publications. --->

@@

We only examined one journal and one trial registry and hence our results may not be generalisable to all journals or registries, especially those that use a statistical reviewer.

>>>>

_PLOS ONE_ is a large open access journal that publishes original research across a wide range of scientific fields. Articles must be in English. Article submissions are handled by an academic editor who selects peer reviewers based on their self-nominated area(s) of expertise. Submissions do not undergo formal statistical review. Instead, reviewers are required to assess submissions against several publication criteria, including whether: “Experiments, statistics, and other analyses are performed to a high technical standard and are described in sufficient detail”. Authors are encouraged to follow published reporting guidelines to ensure that chosen statistical methods are appropriate for the study design, and adequate details are provided to enable independent replication of results. Reporting guidelines are available from the EQUATOR network which has developed guidelines for study designs commonly used in health research, and these guidelines were developed to tackle poor statistical application and reporting in health and medical journals [@Altman2016].

@@

_Step 1_: Targeted API searches. API searches were completed using the R package ‘rplos’ [@rplos]. Search queries targeted the presence of analysis-related terms anywhere in the article. Search terms combined the words “data” or “statistical” with one of: “analysis”, “analyses”, “method”, “methodology” or “model(l)ing”. Search terms were intended to be broad whilst keeping search results to a manageable number for full-text review (see Step 2). By allowing terms to appear anywhere within the article, we accounted for the possibility of relevant text being placed in different sections, for example, in the _Material and Methods_ section versus _Results_. Search results were indexed by a unique Digital Object Identifier (DOI). Attribute data collected per DOI included journal volume, subject classification(s) and total article views since publication.

@@

The ANZCTR was established in 2005 as part of a coordinated global effort to improve research quality and transparency in clinical trials reporting; observational studies can also be registered. All studies registered on ANZCTR are publicly available and can be searched for via an online portal. 
<!--- Researchers responsible for overseeing a clinical trial are expected to register full details before commencing participant enrolment. --->
Details required for registration follow a standardised template (reference or supp file), which covers participant eligibility, intervention(s) being evaluated, study design and outcomes. Researchers are asked to provide a "brief description" of the sample size calculations, statistical methods and planned analyses, although this section is not compulsory [@ANZCTR]. All information must be in English. Studies are reviewed by ANZCTR staff, but are not peer reviewed. 
<!---Upon successful registration, the trial is assigned a unique registration number for use in future reporting in scientific publications. --->

@@

We only examined one journal and one trial registry and hence our results may not be generalisable to all journals or registries, especially those that consistently use a statistical reviewer.

We searched the full text of _PLOS ONE_ papers but not the supporting information which may contain statistical methods sections for some papers. The search terms we used to find statistical methods appeared in the supporting information titles for xxx papers (x%). We did not include the supporting information because it is less structured than the paper and could be in PDF or Word format.
